---
title: "Typing analysis"
author: "Nicholaus Brosowsky"
#output: html_document
date: "08/08/2018"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
pack<-c("dplyr","ggplot2","knitr","kableExtra","car","afex","tidyr")
lapply(pack, require, character.only = TRUE)
```

- Instead of messing around with the livecode stuff, I re-processed all the data from the raw data files (see "to-data-frame.R" & "cleaning-functions.R")
- There are now clean and organized raw data files for each experiment (similar to the machine-learning data frames we used)
- I used new functions to identify words and error words (by comparing each typed word to a list of words from the paragraph)
- This leads to some small differences in the output. The main results however, are all the same (I'm more confident in this analysis than in my 2015 analysis)


```{r}
load(file="E1A/data.Rda")
```

# E1A: speaking letters vs. words (MTurk)

- Note: I ended up with 44 subs here (as opposed to 40). I re-processed from the raw data, so the answer probably lies there. I can't see any issues with the 44 I have, but maybe warrants a closer look.


### Mean word and mean letter error rates
- The only experiment that has letter-level accuracy

```{r}
error.data<-data%>%
  filter(word_letterPos == 1,
         word_length != 0) %>%
  group_by(subject, condition) %>%
  summarize(
    N = n(),
    ER = 1-mean(word_ACC),
    sd = sd(1-word_ACC),
    se = sd / sqrt(N)
  )
  

aov.out <- aov_car(ER ~ condition + Error(subject/condition), data = error.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")

error.summary <- error.data %>%
  group_by(condition) %>%
  summarize(
    N = n(),
    error = mean(ER),
    sd = sd(ER),
    se = sd / sqrt(N)
  )
  

limits <- aes(ymax = error + se, ymin = error - se)
ggplot(error.summary,aes(x=condition, y=error,fill=condition))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0, .25))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("% Word Error Rates")+
  xlab("Innervoice (letters vs. words)")+
  theme(legend.position="top",legend.direction="horizontal")


data$ACC<-as.numeric(as.character(data$ACC))

error.data<-data %>%
  group_by(subject, condition) %>%
  summarize(
    N = n(),
    ER = 1-mean(ACC),
    sd = sd(1-ACC),
    se = sd / sqrt(N)
  )
  

aov.out <- aov_car(ER ~ condition + Error(subject/condition), data = error.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")

error.summary <- error.data %>%
  group_by(condition) %>%
  summarize(
    N = n(),
    error = mean(ER),
    sd = sd(ER),
    se = sd / sqrt(N)
  )
  

limits <- aes(ymax = error + se, ymin = error - se)
ggplot(error.summary,aes(x=condition, y=error,fill=condition))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0, .25))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("% Letter Error Rates")+
  xlab("Innervoice (letters vs. words)")+
  theme(legend.position="top",legend.direction="horizontal")


```

### IKSI analysis

```{r}
iksi.data<-data%>%
  filter(keyNum != 1) %>%
  group_by(subject,condition) %>%
  summarize(meanIKSI = mean(IKSI))


aov.out <- aov_car(meanIKSI ~ condition + Error(subject/condition), data = iksi.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")


iksi.summary<-iksi.data%>%
  group_by(condition)%>%
  summarize(
    N = n(),
    IKSI = mean(meanIKSI),
    sd = sd(meanIKSI),
    se = sd / sqrt(N)
  )

iksi.summary = iksi.summary %>% mutate(condition = reorder(condition, IKSI, mean))

limits <- aes(ymax = IKSI + se, ymin = IKSI - se)
ggplot(iksi.summary,aes(x=condition, y=IKSI,fill=condition))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0, 500))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("IKSI (ms)")+
  xlab("Innervoice (letters vs. words)")+
  theme(legend.position="top",legend.direction="horizontal")

```

# E1B: talking (inner, outer) X type (letters, words)

```{r}
load(file="E1B/data.Rda")
```

### Mean word error rates

```{r}
# 1 sub removed with high error rate (ASDF sub)

error.data<-data%>%
  filter(word_letterPos == 1,
         word_length != 0,
         condition != "Normal",
         subject != "p4c4") %>%
  group_by(subject, talk,type) %>%
  summarize(
    N = n(),
    ER = 1-mean(word_ACC),
    sd = sd(1-word_ACC),
    se = sd / sqrt(N)
  )
  

aov.out <- aov_car(ER ~ talk*type + Error(subject/talk*type), data = error.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")

error.summary <- error.data %>%
  group_by(talk,type) %>%
  summarize(
    N = n(),
    error = mean(ER),
    sd = sd(ER),
    se = sd / sqrt(N)
  )
  

limits <- aes(ymax = error + se, ymin = error - se)
ggplot(error.summary,aes(x=talk, y=error,fill=type))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  #coord_cartesian(ylim = c(0.1, .25))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("% Word Error Rates")+
  xlab("Talking (inner,outer) and type (letters, words)")+
  theme(legend.position="top",legend.direction="horizontal")


```

### Mean IKSI

```{r}
iksi.data<-data%>%
  filter(keyNum != 1,
         subject != "p4c4",
         condition != "Normal") %>%
  group_by(subject,talk,type) %>%
  summarize(meanIKSI = mean(IKSI))


aov.out <- aov_car(meanIKSI ~ talk*type + Error(subject/talk*type), data = iksi.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")


iksi.summary<-iksi.data%>%
  group_by(talk,type)%>%
  summarize(
    N = n(),
    IKSI = mean(meanIKSI),
    sd = sd(meanIKSI),
    se = sd / sqrt(N)
  )


limits <- aes(ymax = IKSI + se, ymin = IKSI - se)
ggplot(iksi.summary,aes(x=talk, y=IKSI,fill=type))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0, 500))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("IKSI (ms)")+
  xlab("Talking (inner,outer) and type (letters, words)")+
  theme(legend.position="top",legend.direction="horizontal")

```

# E2: Verbal Suppression

```{r}
load(file="E2/data.Rda")
```

### Mean word error rates


```{r}
error.data<-data%>%
  filter(word_letterPos == 1,
         word_length != 0) %>%
  group_by(subject, condition) %>%
  summarize(
    N = n(),
    ER = 1-mean(word_ACC),
    sd = sd(1-word_ACC),
    se = sd / sqrt(N)
  )
  

aov.out <- aov_car(ER ~ condition + Error(subject/condition), data = error.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")

error.summary <- error.data %>%
  group_by(condition) %>%
  summarize(
    N = n(),
    error = mean(ER),
    sd = sd(ER),
    se = sd / sqrt(N)
  )

error.summary = error.summary %>% mutate(condition = reorder(condition, error, mean))


limits <- aes(ymax = error + se, ymin = error - se)
ggplot(error.summary,aes(x=condition, y=error,fill=condition))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0.1, .25))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("Error Rates")+
  xlab("Verbal suppression")+
  theme(legend.position="top",legend.direction="horizontal")
```




```{r}
iksi.data<-data%>%
  filter(keyNum != 1) %>%
  group_by(subject,condition) %>%
  summarize(meanIKSI = mean(IKSI))


aov.out <- aov_car(meanIKSI ~ condition + Error(subject/condition), data = iksi.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")


iksi.summary<-iksi.data%>%
  group_by(condition)%>%
  summarize(
    N = n(),
    IKSI = mean(meanIKSI),
    sd = sd(meanIKSI),
    se = sd / sqrt(N)
  )

iksi.summary = iksi.summary %>% mutate(condition = reorder(condition, IKSI, mean))

limits <- aes(ymax = IKSI + se, ymin = IKSI - se)
ggplot(iksi.summary,aes(x=condition, y=IKSI,fill=condition))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0, 500))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("IKSI (ms)")+
  xlab("Verbal suppression")+
  theme(legend.position="top",legend.direction="horizontal")


```

# E3: Delayed auditory feedback


```{r}
load(file="E3/data.Rda")
```


### Mean word error rates


```{r}
error.data<-data%>%
  filter(word_letterPos == 1,
         word_length != 0) %>%
  group_by(subject, delay) %>%
  summarize(
    N = n(),
    ER = 1-mean(word_ACC),
    sd = sd(1-word_ACC),
    se = sd / sqrt(N)
  )
  

aov.out <- aov_car(ER ~ delay + Error(subject/delay), data = error.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")

error.summary <- error.data %>%
  group_by(delay) %>%
  summarize(
    N = n(),
    error = mean(ER),
    sd = sd(ER),
    se = sd / sqrt(N)
  )
  

limits <- aes(ymax = error + se, ymin = error - se)
ggplot(error.summary,aes(x=delay, y=error,fill=delay))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(0.1, .25))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("Error Rates")+
  xlab("Delayed Feedback")+
  theme(legend.position="top",legend.direction="horizontal")
```

### Mean IKSI

```{r}
iksi.data<-data%>%
  filter(keyNum != 1) %>%
  group_by(subject,delay) %>%
  summarize(meanIKSI = mean(IKSI))


aov.out <- aov_car(meanIKSI ~ delay + Error(subject/delay), data = iksi.data)

kable(nice(aov.out, es = "pes", observed = NULL,
    correction = c("none"), MSE = TRUE,
    sig.symbols = c(" +", " *", " **", " ***"))
,caption = "ANOVA")


iksi.summary<-iksi.data%>%
  group_by(delay)%>%
  summarize(
    N = n(),
    IKSI = mean(meanIKSI),
    sd = sd(meanIKSI),
    se = sd / sqrt(N)
  )
  

limits <- aes(ymax = IKSI + se, ymin = IKSI - se)
ggplot(iksi.summary,aes(x=delay, y=IKSI,fill=delay))+
  geom_bar(stat="identity", position=position_dodge(width=0.9),color="black") + 
  geom_errorbar(limits, width = .2, position=position_dodge(width=0.9))+
  coord_cartesian(ylim = c(200, 325))+
  theme_classic(base_size=12)+
  labs(title="")+
  ylab("IKSI (ms)")+
  xlab("Delayed Feedback")+
  theme(legend.position="top",legend.direction="horizontal")


```

