---
title: "Analysis E2 Verbal Suppression"
author: "Matt"
date: "5/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = TRUE)
```

# Load Libraries

```{r}
library(papaja)
library(TalkTyping)
library(dplyr)
library(Crump)
library(xtable)
library(ggplot2)
library(ggpubr)
library(forcats)
```

# E2 IKSI Analysis

```{r}
#load E2 data
E2_data <- talk_type_E2_data

# IKSI analysis

E2_data <- E2_data %>%
           filter(letter_accuracy == 1,
                           iksis < 5000,
                           LetterType != "Space") %>%
            mutate(subject = as.factor(subject),
                   suppression = as.factor(suppression),
                   LetterType = as.factor(LetterType)) %>%
            group_by(subject,suppression,LetterType) %>%
            summarise(mean_iksi = mean(modified_recursive_moving(iksis)$restricted),
                      prop_removed = modified_recursive_moving(iksis)$prop_removed)
                  
E2_aov_out <- aov(mean_iksi ~ suppression*LetterType + 
                    Error(subject/(suppression*LetterType)), E2_data)

knitr::kable(xtable(summary(E2_aov_out)))

E2_apa_print <- apa_print(E2_aov_out)
E2_means <- model.tables(E2_aov_out,"means")

```

# First Letter one-way with linear contrasts

```{r}
#load E1B data
E2_data <- talk_type_E2_data

# IKSI analysis

E2_data$suppression <- fct_relevel(E2_data$suppression,c("Normal","SayThe","TueThur","Alphabet","RandLetter","Count"))

E2_data <- E2_data %>%
           filter(letter_accuracy == 1,
                           iksis < 5000,
                           LetterType != "Space") %>%
            mutate(subject = as.factor(subject),
                   suppression = as.factor(suppression),
                   LetterType = as.factor(LetterType)) %>%
            group_by(subject,suppression,LetterType) %>%
            summarise(mean_iksi = mean(modified_recursive_moving(iksis)$restricted),
                      prop_removed = modified_recursive_moving(iksis)$prop_removed)

E2_FL_data <- E2_data %>%
                filter(LetterType == "First")

E2_FL_aov_out <- aov(mean_iksi ~ suppression + 
                    Error(subject/(suppression)), E2_FL_data)

knitr::kable(xtable(summary(E2_FL_aov_out)))

E2_FL_apa_print <- apa_print(E2_FL_aov_out)
E2_FL_means <- model.tables(E2_FL_aov_out,"means")

```

## Plot

```{r}
E2_FL_iksi_table <- E2_FL_data %>%
              group_by(suppression) %>%
              summarize(mIKSI = mean(mean_iksi),
                        sem = sd(mean_iksi)/sqrt(length(mean_iksi)))

E2_FL_graph_iksi <- ggplot(E2_FL_iksi_table, aes(x=suppression,
                                             y=mIKSI))+
  geom_bar(stat="identity", position="dodge", fill="grey")+
  geom_errorbar(aes(ymin=mIKSI-sem,
                    ymax=mIKSI+sem), width=.1,
                linetype="solid", position=position_dodge(.9))+
  scale_fill_grey(start = 0.6, end = 0.8, na.value = "red",
  aesthetics = "fill")+
  theme_classic(base_size=9)+
  theme(legend.position = "top",
        legend.title = element_blank())+
  coord_cartesian(ylim=c(100,800))+
  scale_y_continuous(minor_breaks=seq(100,800,25),
                     breaks=seq(100,800,100))+
   #scale_x_discrete(labels = c('Normal',
  #                            'Letter',
  #                            'Word',
  #                            'Letter',
  #                            'Word'))+
  ylab("Mean IKSI (ms)")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 # facet_wrap(~suppression, scales = "free_x",
 #            strip.position="bottom")

knitr::kable(E2_FL_iksi_table)
E2_FL_graph_iksi
```


## contrasts

```{r}
# Normal vs. All
E2_FL_NvsAll <- apa_print(t_contrast_rm(df=E2_FL_data,
                             subject = "subject",
                             dv = "mean_iksi",
                             condition = "suppression",
                             A_levels = c("Normal"),
                             B_levels = c("SayThe","TueThur","Alphabet","RandLetter","Count"),
                             contrast_weights = c(-1,1/5,1/5,1/5,1/5,1/5) ))

# Normal vs. the and tuesday thursday
E2_FL_NvsTheThu <- apa_print(t_contrast_rm(df=E2_FL_data,
                             subject = "subject",
                             dv = "mean_iksi",
                             condition = "suppression",
                             A_levels = c("Normal"),
                             B_levels = c("SayThe","TueThur"),
                             contrast_weights = c(-1,1/2,1/2) ))

#  the and tuesday thursday vs. alphabet
E2_FL_AlphavsTheThu <- apa_print(t_contrast_rm(df=E2_FL_data,
                             subject = "subject",
                             dv = "mean_iksi",
                             condition = "suppression",
                             A_levels = c("SayThe","TueThur"),
                             B_levels = c("Alphabet"),
                             contrast_weights = c(-1/2,-1/2,1) ))

E2_FL_AlphavsRandCnt <- apa_print(t_contrast_rm(df=E2_FL_data,
                             subject = "subject",
                             dv = "mean_iksi",
                             condition = "suppression",
                             A_levels = c("Alphabet"),
                             B_levels = c("RandLetter","Count"),
                             contrast_weights = c(-1,1/2,1/2) ))
```


# Middle Letter one-way with linear contrasts

```{r}
#load E1B data
E2_data <- talk_type_E2_data

# IKSI analysis

E2_data$suppression <- fct_relevel(E2_data$suppression,c("Normal","SayThe","TueThur","Alphabet","RandLetter","Count"))

E2_data <- E2_data %>%
           filter(letter_accuracy == 1,
                           iksis < 5000,
                           LetterType != "Space") %>%
            mutate(subject = as.factor(subject),
                   suppression = as.factor(suppression),
                   LetterType = as.factor(LetterType)) %>%
            group_by(subject,suppression,LetterType) %>%
            summarise(mean_iksi = mean(modified_recursive_moving(iksis)$restricted),
                      prop_removed = modified_recursive_moving(iksis)$prop_removed)

E2_ML_data <- E2_data %>%
                filter(LetterType == "Middle")

E2_ML_aov_out <- aov(mean_iksi ~ suppression + 
                    Error(subject/(suppression)), E2_ML_data)

knitr::kable(xtable(summary(E2_ML_aov_out)))

E2_ML_apa_print <- apa_print(E2_ML_aov_out)
E2_ML_means <- model.tables(E2_ML_aov_out,"means")

```

## Plot

```{r}
E2_ML_iksi_table <- E2_ML_data %>%
              group_by(suppression) %>%
              summarize(mIKSI = mean(mean_iksi),
                        sem = sd(mean_iksi)/sqrt(length(mean_iksi)))

E2_ML_graph_iksi <- ggplot(E2_ML_iksi_table, aes(x=suppression,
                                             y=mIKSI))+
  geom_bar(stat="identity", position="dodge", fill="grey")+
  geom_errorbar(aes(ymin=mIKSI-sem,
                    ymax=mIKSI+sem), width=.1,
                linetype="solid", position=position_dodge(.9))+
  scale_fill_grey(start = 0.6, end = 0.8, na.value = "red",
  aesthetics = "fill")+
  theme_classic(base_size=9)+
  theme(legend.position = "top",
        legend.title = element_blank())+
  coord_cartesian(ylim=c(100,200))+
  scale_y_continuous(minor_breaks=seq(100,200,25),
                     breaks=seq(100,200,25))+
   #scale_x_discrete(labels = c('Normal',
  #                            'Letter',
  #                            'Word',
  #                            'Letter',
  #                            'Word'))+
  ylab("Mean IKSI (ms)")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 # facet_wrap(~suppression, scales = "free_x",
 #            strip.position="bottom")

knitr::kable(E2_ML_iksi_table)
E2_ML_graph_iksi
```


## contrasts

```{r}
# Normal vs. all but random

E2_ML_NvsAllbR <- apa_print(t_contrast_rm(df=E2_ML_data,
                             subject = "subject",
                             dv = "mean_iksi",
                             condition = "suppression",
                             A_levels = c("Normal"),
                             B_levels = c("SayThe","TueThur","Alphabet","Count"),
                             contrast_weights = c(-1,1/4,1/4,1/4,1/4) ))

# Rand vs. Normal
E2_ML_RandvsNormal <- apa_print(t_contrast_rm(df=E2_ML_data,
                             subject = "subject",
                             dv = "mean_iksi",
                             condition = "suppression",
                             A_levels = c("Normal"),
                             B_levels = c("RandLetter"),
                             contrast_weights = c(-1,1) ))

```



# First Letter Accuracy one-way with linear contrasts

```{r}
#load E1B data
E2_data <- talk_type_E2_data

# IKSI analysis

E2_data$suppression <- fct_relevel(E2_data$suppression,c("Normal","SayThe","TueThur","Alphabet","RandLetter","Count"))

E2_data <- E2_data %>%
           filter(LetterType != "Space") %>%
            mutate(subject = as.factor(subject),
                   suppression = as.factor(suppression),
                   LetterType = as.factor(LetterType)) %>%
            group_by(subject,suppression,LetterType) %>%
            summarise(mean_acc = mean(letter_accuracy))

E2acc_FL_data <- E2_data %>%
                filter(LetterType == "First")

E2acc_FL_aov_out <- aov(mean_acc ~ suppression + 
                    Error(subject/(suppression)), E2acc_FL_data)

knitr::kable(xtable(summary(E2acc_FL_aov_out)))

E2acc_FL_apa_print <- apa_print(E2acc_FL_aov_out)
E2acc_FL_means <- model.tables(E2acc_FL_aov_out,"means")

```

## Plot

```{r}
E2acc_FL_table <- E2acc_FL_data %>%
              group_by(suppression) %>%
              summarize(mAcc = mean(mean_acc),
                        sem = sd(mean_acc)/sqrt(length(mean_acc)))

E2acc_FL_graph_acc <- ggplot(E2acc_FL_table, aes(x=suppression,
                                             y=mAcc))+
  geom_bar(stat="identity", position="dodge", fill="grey")+
  geom_errorbar(aes(ymin=mAcc-sem,
                    ymax=mAcc+sem), width=.1,
                linetype="solid", position=position_dodge(.9))+
  scale_fill_grey(start = 0.6, end = 0.8, na.value = "red",
  aesthetics = "fill")+
  theme_classic(base_size=9)+
  theme(legend.position = "top",
        legend.title = element_blank())+
  coord_cartesian(ylim=c(0.7,1))+
  scale_y_continuous(minor_breaks=seq(0.7,1,.1),
                     breaks=seq(0.7,1,.1))+
   #scale_x_discrete(labels = c('Normal',
  #                            'Letter',
  #                            'Word',
  #                            'Letter',
  #                            'Word'))+
  ylab("Mean Accuracy")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 # facet_wrap(~suppression, scales = "free_x",
 #            strip.position="bottom")

knitr::kable(E2acc_FL_table)
E2acc_FL_graph_acc
```


## contrasts

```{r}
# Normal vs. the tue/thur

E2acc_FL_NvsTheThur <- apa_print(t_contrast_rm(df=E2acc_FL_data,
                             subject = "subject",
                             dv = "mean_acc",
                             condition = "suppression",
                             A_levels = c("Normal"),
                             B_levels = c("SayThe","TueThur"),
                             contrast_weights = c(-1,1/2,1/2) ))

#  Normal the tue/thur vs alpha rand count
E2acc_FL_NTheThvsARC <- apa_print(t_contrast_rm(df=E2acc_FL_data,
                             subject = "subject",
                             dv = "mean_acc",
                             condition = "suppression",
                             A_levels = c("Normal","SayThe","TueThur"),
                             B_levels = c("Alphabet","RandLetter","Count"),
                             contrast_weights = c(-1/3,-1/3,-1/3,1/3,1/3,1/3) ))

```



# Middle Letter Accuracy one-way with linear contrasts

```{r}
#load E1B data
E2_data <- talk_type_E2_data

# IKSI analysis

E2_data$suppression <- fct_relevel(E2_data$suppression,c("Normal","SayThe","TueThur","Alphabet","RandLetter","Count"))

E2_data <- E2_data %>%
           filter(LetterType != "Space") %>%
            mutate(subject = as.factor(subject),
                   suppression = as.factor(suppression),
                   LetterType = as.factor(LetterType)) %>%
            group_by(subject,suppression,LetterType) %>%
            summarise(mean_acc = mean(letter_accuracy))

E2acc_ML_data <- E2_data %>%
                filter(LetterType == "Middle")

E2acc_ML_aov_out <- aov(mean_acc ~ suppression + 
                    Error(subject/(suppression)), E2acc_ML_data)

knitr::kable(xtable(summary(E2acc_ML_aov_out)))

E2acc_ML_apa_print <- apa_print(E2acc_ML_aov_out)
E2acc_ML_means <- model.tables(E2acc_ML_aov_out,"means")

```

## Plot

```{r}
E2acc_ML_table <- E2acc_ML_data %>%
              group_by(suppression) %>%
              summarize(mAcc = mean(mean_acc),
                        sem = sd(mean_acc)/sqrt(length(mean_acc)))

E2acc_ML_graph_acc <- ggplot(E2acc_ML_table, aes(x=suppression,
                                             y=mAcc))+
  geom_bar(stat="identity", position="dodge", fill="grey")+
  geom_errorbar(aes(ymin=mAcc-sem,
                    ymax=mAcc+sem), width=.1,
                linetype="solid", position=position_dodge(.9))+
  scale_fill_grey(start = 0.6, end = 0.8, na.value = "red",
  aesthetics = "fill")+
  theme_classic(base_size=9)+
  theme(legend.position = "top",
        legend.title = element_blank())+
  coord_cartesian(ylim=c(0.7,1))+
  scale_y_continuous(minor_breaks=seq(0.7,1,.1),
                     breaks=seq(0.7,1,.1))+
   #scale_x_discrete(labels = c('Normal',
  #                            'Letter',
  #                            'Word',
  #                            'Letter',
  #                            'Word'))+
  ylab("Mean Accuracy")+
  theme(axis.title.x = element_blank()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 # facet_wrap(~suppression, scales = "free_x",
 #            strip.position="bottom")

knitr::kable(E2acc_ML_table)
E2acc_ML_graph_acc
```


## contrasts

```{r}
# Normal vs. the tue/thur

E2acc_ML_NvsTheThur <- apa_print(t_contrast_rm(df=E2acc_ML_data,
                             subject = "subject",
                             dv = "mean_acc",
                             condition = "suppression",
                             A_levels = c("Normal"),
                             B_levels = c("SayThe","TueThur"),
                             contrast_weights = c(-1,1/2,1/2) ))

#  Normal the tue/thur vs alpha rand count
E2acc_ML_NTheThvsARC <- apa_print(t_contrast_rm(df=E2acc_ML_data,
                             subject = "subject",
                             dv = "mean_acc",
                             condition = "suppression",
                             A_levels = c("Normal","SayThe","TueThur"),
                             B_levels = c("Alphabet","RandLetter","Count"),
                             contrast_weights = c(-1/3,-1/3,-1/3,1/3,1/3,1/3) ))

```


# speed accuracy tradeoff

```{r}
library(ggrepel)
E2_iksi_both <- rbind(E2_FL_iksi_table,
                       E2_ML_iksi_table)

E2_iksi_both <- cbind(E2_iksi_both, 
                       Letter_Position = rep(c("First","Middle"), each=6))

E2_acc_both <- rbind(E2acc_FL_table,
                       E2acc_ML_table)

E2_acc_both <- cbind(E2_acc_both, 
                       Letter_Position = rep(c("First","Middle"), each=6))

E2_SA <- cbind(E2_iksi_both, accuracy = E2_acc_both$mAcc)

E2_SA_graph <- ggplot(E2_SA, aes(x=mIKSI,  y=accuracy,
                   shape=Letter_Position,
                   color=Letter_Position,
                   label=suppression))+
  geom_point()+
  geom_text_repel(size=1.7, color="black")+
  coord_cartesian(xlim=c(100,700), ylim=c(.8,1))+
  scale_x_continuous(breaks=seq(100,700,100))+
  scale_color_grey(start = 0.2, end = 0.5, na.value = "red",
  aesthetics = "color", guide =FALSE)+
  theme_classic(base_size=10)+
  ylab("Mean Accuracy")+
  xlab("Mean IKSI (ms)")+
  #facet_wrap(~Letter_Position, nrow=2,
  #           strip.position="right")+
  theme(legend.position ="top",
        legend.direction = "horizontal")+
  guides(shape = guide_legend(label.hjust = 0,
                              keywidth=0.1))+
  labs(shape="Letter")
  

E2_SA_graph

```

```{r, fig.width=6, fig.height=3}
#load E2 data
E2_data <- talk_type_E2_data

# IKSI analysis

E2_data <- E2_data %>%
           filter(accuracy == 1,
                           iksis < 5000,
                           LetterType != "Space") %>%
            mutate(subject = as.factor(subject),
                   suppression = as.factor(suppression),
                   LetterType = as.factor(LetterType)) %>%
            group_by(subject,suppression,LetterType) %>%
            summarise(mean_iksi = mean(modified_recursive_moving(iksis)$restricted),
                      prop_removed = modified_recursive_moving(iksis)$prop_removed)

E2_data$suppression <- fct_relevel(E2_data$suppression,c("Normal","SayThe","TueThur","Alphabet","RandLetter","Count"))

E2_iksi_table <- E2_data %>%
              group_by(suppression,LetterType) %>%
              summarize(mIKSI = mean(mean_iksi),
                        sem = sd(mean_iksi)/sqrt(length(mean_iksi)))

#levels(E1A_iksi_table$paragraph) <- c("Say Letter", "Say Word")

E2_graph_iksi <- ggplot(E2_iksi_table, aes(x=LetterType,
                                             y=mIKSI,
                                           fill=LetterType))+
  geom_bar(stat="identity",position="dodge")+
  geom_errorbar(aes(ymin=mIKSI-sem,
                    ymax=mIKSI+sem), width=.1,
                linetype="solid")+
  scale_fill_grey(start = 0.6, end = 0.8, na.value = "red",
  aesthetics = "fill")+
  theme_classic(base_size=12)+
  theme(legend.position = "none",
        legend.title = element_blank())+
  ylab("Mean IKSI (ms)")+
  xlab("Letter position")+
  facet_wrap(~suppression, nrow=1)

knitr::kable(E2_iksi_table)
E2_graph_iksi

```

## accuracy

## get letter level accuracy estimates

This chunk was previously evaluated, and the `talk_type_e2_data.RData` file was updated with estimates of letter level accuracy.

```{r, eval=FALSE}

E2acc_data <- talk_type_E2_data

E2acc_data$closest[is.na(E2acc_data$closest)] <- ""

letter_accuracy <- E2acc_data$accuracy
for(i in 1:dim(E2acc_data)[1]){
  clet <- as.character(E2acc_data[i,]$letters)
  cword <- unlist(strsplit(as.character(E2acc_data[i,]$closest),split=""))
  if(E2acc_data$accuracy[i] == 0){
    if(length(cword)>0){
      if(E2acc_data[i,]$LetterPosition <= length(cword)){
        if(clet != " "){
          if(clet == cword[E2acc_data[i,]$LetterPosition]){
            letter_accuracy[i] <- 1
          } else {
            letter_accuracy[i] <- 0
          }
        }
      }
    }
  }
}

E2acc_data <- cbind(E2acc_data,letter_accuracy)
```

## Accuracy ANOVA

```{r}
E2acc_data <- talk_type_E2_data

# Accuracy

E2acc_data <- E2acc_data %>%
                mutate(subject = as.factor(subject),
                       suppression = as.factor(suppression),
                       LetterType = as.factor(LetterType)) %>%
                filter(LetterType != "Space") %>%
                group_by(subject,suppression, LetterType) %>%
                summarise(mean_acc = mean(letter_accuracy))
                  
E2acc_aov_out <- aov(mean_acc ~ suppression*LetterType + Error(subject/(suppression*LetterType)), E2acc_data)

E2acc_apa_print <- apa_print(E2acc_aov_out)
E2acc_means <- model.tables(E2acc_aov_out,"means")

knitr::kable(xtable(summary(E2acc_aov_out)))

```

## Accuracy table and plots


```{r, fig.width=6, fig.height=3}

E2acc_data$suppression <- fct_relevel(E2acc_data$suppression,c("Normal","SayThe","TueThur","Alphabet","RandLetter","Count"))

E2_acc_table <- E2acc_data %>%
              group_by(suppression,LetterType) %>%
              summarize(mAcc = mean(mean_acc),
                        sem = sd(mean_acc)/sqrt(length(mean_acc)))

#levels(E1A_acc_table$paragraph) <- c("Say Letter", "Say Word")

E2_graph_acc <- ggplot(E2_acc_table, aes(x=LetterType,
                                             y=mAcc,
                                           fill=LetterType))+
  geom_bar(stat="identity",position="dodge")+
  geom_errorbar(aes(ymin=mAcc-sem,
                    ymax=mAcc+sem), width=.1,
                linetype="solid")+
  scale_fill_grey(start = 0.6, end = 0.8, na.value = "red",
  aesthetics = "fill")+
  theme_classic(base_size=12)+
  theme(legend.position = "none",
        legend.title = element_blank())+
  ylab("Mean Accuracy")+
  xlab("Letter position")+
  coord_cartesian(ylim=c(.7,1))+
  facet_wrap(~suppression, nrow=1)

knitr::kable(E2_acc_table)
E2_graph_acc
```

# Figure 1

```{r, eval=T, fig.width=6, fig.height=6}
ggarrange(E2_graph_iksi,E2_graph_acc, nrow=2)

ggarrange(E2_FL_graph_iksi,E2_ML_graph_iksi, nrow=2)

```

# Speed accuracy tradeoff

```{r}
tradeoff <- E2_iksi_table
tradeoff <- cbind(tradeoff,accuracy = E2_acc_table$mAcc,
                  accuracy_sem = E2_acc_table$sem)

ggplot(tradeoff, aes(x=mIKSI, y=accuracy, 
                     group = suppression,
                     shape= LetterType,
                     linetype = suppression))+
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin=accuracy-accuracy_sem,
                    ymax=accuracy+accuracy_sem), width=.1,
                linetype="solid")+
  geom_errorbarh(aes(xmin=mIKSI-sem,
                    xmax=mIKSI+sem), width=.2,
                linetype="solid")+
  theme_classic(base_size=12)+
  xlab("Mean IKSI (ms)")+
  ylab("Mean Accuracy")
```


# Save all

```{r, eval=T}
save.image(file="E2_workspace.RData")
```


```{r, eval=FALSE}
talk_type_E1B_data <- talk_type_E1B_data %>%
                        rename(voice = VoiceCondition,
                               linguistic_unit = ChunkCondition)

save(talk_type_E1B_data, file="talk_type_E1B_data.RData")
```

